\huge \textbf{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}

### 2025-04-22 paper/paper_draft.tex
paper changes

To constrain cosmological parameters we can then look at the lensing potential of the CMB, $\psi_\CMB$, and of galaxy surveys, $\psi_\gal$. These are directly related to the matter power and bispectra as
\begin{empheq}[box=\fbox]{align*}
    P^{\psi_X\psi_Y}_l
    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(l/\chi,\eta_0-\chi),\\
    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi),
\end{empheq}
with $X, Y, Z \in \{\CMB, \gal\}$. $\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $P^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
$$
\begin{pmatrix}
    l_1&l_2&l_3 \\ m_1 & m_2 & m_3
\end{pmatrix}
$$ 
the Wigner 3-j symbol. The only factor that differs between CMB and galaxy lensing is the window function, defined as
\begin{equation*}
    W_X(\chi) = \int_\chi^{\infty}\d\chi' p_X(\chi')\frac{\chi'-\chi}{\chi'\chi},
\end{equation*}
with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$. Notice that we can also look at cross correlation between CMB and galaxy lensing.

@@

\subsection{Nonlinear matter bispectrum}
Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB. The nonlinear matter bispectrum was calculated from the powerspectrum using a fitting formula based on perturbation theory in \cite{bispfit}. It is given by
\begin{equation}
    B_m(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P_m(k_1, z) P_m(k_2, z) + \text{2 perm.}
\end{equation}
where \( P_m \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
\begin{equation}
    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
\end{equation}
They are defined as:
\begin{equation}
    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
\end{equation}
\begin{equation}
    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
\end{equation}
\begin{equation}
    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
\end{equation}
Here, \( Q(n_{\text{eff}}) \) is given by:
\begin{equation}
    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
\end{equation}
The effective spectral index of the linear power spectrum is defined as:
\begin{equation}
    n_{\text{eff}} \equiv \frac{d \ln P_m^{\text{lin}}(k)}{d \ln k}.
\end{equation}
Additionally, \( q \) is given by:
\begin{equation}
    q = \frac{k}{k_{\text{NL}}},
\end{equation}
where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
\begin{equation}
    4 \pi k_{\text{NL}}^3 P_m^{\text{lin}}(k_{\text{NL}}, 0) = 1.
\end{equation}
The coefficients \( a_i \) are:
\begin{align*}
    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
\end{align*}

@@

\section{Weak Lensing}\label{sec:weaklensing}
\subsection{Perturbed Photon Paths}
We work in the conformal newtonian gauge and with natural units. Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element in FLRW spacetime is given by
\begin{equation}
    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j),
\end{equation}
where $\gamma_{ij}$ is the unperturbed line element
\begin{equation}
    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2),
\end{equation}
and $f_K(\chi)$ is the comoving angular diameter distance. We will hereafter only consider a flat universe, so that $f_K(\chi) = \chi$. Weak lensing of a point source can be quantified by looking at the deflection field $\mathbf d (\hat{\mathbf n}) = \mathbf \theta_{\text{obs}} - \mathbf \theta_{\text{true}}$, i.e. the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occured. To first order in $\Psi_N$ and $\Phi_N$, the deflection is given as \cite{dodelson2020modern}
\begin{equation}
    \mathbf d(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi),
\end{equation}
where $\Psi$ is the Weyl Potential, $\Psi := (\Psi_N - \Phi_N)/2$, and $\chi_*$ is the conformal distance to the source. $\nabla_{\hat{\mathbf n}}$ is the derivative along the axes orthogonal to the line of sight. The above equation can be written in terms of the lensing potential, $\psi$, as $\mathbf d (\hat{\mathbf n}) = \nabla_{\hat{\mathbf n}}$, with
\begin{equation}
    \psi(\hat {\mathbf n}) := -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
\end{equation}

If the source is instead distributed radial distances according to some distribution function $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, the $(\chi_* - \chi)/(\chi_*\chi)$ factor is changed as
\begin{equation*}
    \frac{\chi_*-\chi}{\chi_*\chi} \rightarrow W(\chi) := \int_\chi^{\infty}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
\end{equation*}
$W(\chi)$ is then called the window function. In the most general case, the lensing potential is thus given by
\begin{equation}
    \psi(\hat {\mathbf n}) := -2 \int_0^{\infty}\d \chi W(\chi)\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
\end{equation}
The integration limit is sometimes also taken to be the surface of last scattering, as any window function vanishes after that distance. In the case of CMB lensing we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces to $H(\chi_* - \chi)(\chi_*-\chi)/(\chi_*\chi)$, with $H(\chi)$ the Heaviside step function.

@@

\subsection{Convergence and Shear}
Quantities we are interested in measuring depend on how the lensing vector changes with the observation angle. The magnification matrix is defined as
\begin{equation}
    A_{ij} := \delta_{ij} + \frac{\partial}{\partial\theta_i}\alpha_i(\hat{\mathbf n}).
\end{equation}
This matrix can then be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
\begin{equation}
    A_ij(\hat{\mathbf n}) = \begin{pmatrix}
        1 - \kappa - \gamma_1 & -\gamma_2 + \omega \\
        -\gamma_2 - \omega & 1 - \kappa + \gamma_1
    \end{pmatrix}.
\end{equation}
At 1st order $A$ is a symetric matrix by definition and so $\omega$ has to be 0, we will ignore it from here on out. Intuitively, A tells you how a small patch in the sky transforms due to weak lensing, i.e. if you have surface brightness $I$ then
% TODO: make delta thick
\begin{equation}
    I_{\text{emitted}}(\hat{\mathbf{n}} + \mathbf{\delta}) = I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\delta}) + \mathbf\alpha(\hat{\mathbf{n}} + \mathbf{\delta})) \approx I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\alpha} + A_{ij}\delta_{j}).
\end{equation}
given that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$, we can interpret $\kappa$ as telling us about the overal magnification of the source. The $\gamma_i$ represent the area preserving distortion, i.e. stretching and squeezing in a specific direction.

We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
% TODO: should these subscripts be theta and phi?
\begin{equation}
    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_\theta^2 - \partial_\phi^2)\psi, \quad \gamma_2 = \partial_\theta\partial_\phi \psi.
\end{equation}

@@

Consider the lensing potential, $\psi$, then decomposing into spherical harmonics gives

@@

On the other hand, consider the decomposition of $\Psi$ in fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,

@@

where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (Born approximation), with $\eta_0$ the time at which the light ray hits earth. 

@@

To further evaluate the integral we will make the standard \textbf{Limber approximation}. The Bessel functions peak sharply at $l=x$\footnote{Some sources use $x\approx l+1/2$ instead, which is slightly more accurate for larger scales (low $l$) and slightly less accurate for smaller scales.}, with the peak being increasingly sharp for higher $l$. Similarly to $\delta(x-x_0)f(x)=\delta(x-x_0)f(x_0)$, we thus take $j_l(k\chi)f(k)\approx j_l(k\chi)f(l/\chi)$. The Bessel functions satisfy an orthogonality condition,

@@

The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is \cite{Hu2000}
\begin{equation*}
    \langle (X_1)_{l_1m_1}(X_2)_{l_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
        l_1 & l_2 & l_3 \\ m_1 & m_2 & m_3
    \end{pmatrix}
    B_{l_1l_2l_3}^{X_1X_2X_3}.
\end{equation*}
Note the independence on $m_i$, this necesarilly follows from statistical isotropy. % TODO: add a short explanation here
If $m_1+m_2+m_3\neq 0$, the associated Wigner-3j symbol vanishes and the bispectrum is set to zero.
Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$. This will be relevant later when we look at cross correlations between shear and convergence fields.

Using the above definition and the symmetry property
\begin{equation*}
    \begin{pmatrix}
        l_1 & l_2 & l_3\\
        -m_1 & -m_2 & -m_3
      \end{pmatrix}
      =
      (-1)^{l_1+l_2+l_3}
      \begin{pmatrix}
        l_1 & l_2 & l_3\\
        m_1 & m_2 & m_3
      \end{pmatrix}
\end{equation*}
we find
\begin{gather*}
    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
    = (-1)^{l_1+l_2+l_3}
    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
    (2\pi)^3 8 \\
    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \{\eta_0-\chi\}),
\end{gather*}
where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispecrum vanishing if that sum doesn't equal $0$, as mentioned earlier. When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
In particular, it vanishes if $l_1+l_2+l_3$ is odd, meaning we can drop the $(-1)^{l_1+l_2+l_3}$ factor. Additionally cancelling common factors then gives
\begin{align}
    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
    =
    -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi). \label{eq:lbsintermsofpsi}
\end{align}

>>>>

To constrain cosmological parameters we can then look at the lensing potential of the CMB, $\psi_\CMB$, and of galaxy surveys, $\psi_\gal$. These are directly related to the matter power and bispectra as
\begin{empheq}[box=\fbox]{align*}
    C^{\psi_X\psi_Y}_l
    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(l/\chi,\eta_0-\chi),\\
    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi),
\end{empheq}
with $X, Y, Z \in \{\CMB, \gal\}$. $\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $P^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
$$
\begin{pmatrix}
    l_1&l_2&l_3 \\ m_1 & m_2 & m_3
\end{pmatrix}
$$ 
the Wigner 3-j symbol. The only factor that differs between CMB and galaxy lensing is the window function, defined as
\begin{equation*}
    W_X(\chi) = \int_\chi^{\infty}\d\chi' p_X(\chi')\frac{\chi'-\chi}{\chi'\chi},
\end{equation*}
with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$. Notice that we can also look at cross correlation between CMB and galaxy lensing.

@@

\subsection{Nonlinear matter bispectrum}
Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB. The nonlinear matter bispectrum was calculated from the powerspectrum using a fitting formula based on perturbation theory in \cite{bispfit}. It is given by
\begin{equation}
    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) C^\delta(k_1, z) C^\delta(k_2, z) + \text{2 perm.}
\end{equation}
where \( C^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
\begin{equation}
    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
\end{equation}
They are defined as:
\begin{equation}
    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
\end{equation}
\begin{equation}
    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
\end{equation}
\begin{equation}
    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
\end{equation}
Here, \( Q(n_{\text{eff}}) \) is given by:
\begin{equation}
    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
\end{equation}
The effective spectral index of the linear power spectrum is defined as:
\begin{equation}
    n_{\text{eff}} \equiv \frac{d \ln C^\delta_{\text{lin}}(k)}{d \ln k}.
\end{equation}
Additionally, \( q \) is given by:
\begin{equation}
    q = \frac{k}{k_{\text{NL}}},
\end{equation}
where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
\begin{equation}
    4 \pi k_{\text{NL}}^3 C^\delta_{\text{lin}}(k_{\text{NL}}, 0) = 1.
\end{equation}
The coefficients \( a_i \) are:
\begin{align*}
    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
\end{align*}

@@

\section{Weak Lensing}\label{sec:weaklensing}
\subsection{Perturbed Photon Paths}
We work in the conformal newtonian gauge and with natural units. Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element in FLRW spacetime is given by
\begin{equation}
    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j),
\end{equation}
where $\gamma_{ij}$ is the unperturbed line element
\begin{equation}
    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2),
\end{equation}
and $f_K(\chi)$ is the comoving angular diameter distance. We will hereafter only consider a flat universe, so that $f_K(\chi) = \chi$. Weak lensing of a point source can be quantified by looking at the deflection field $\mathbf d (\hat{\mathbf n}) = \mathbf \theta_{\text{obs}} - \mathbf \theta_{\text{true}}$, i.e. the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occured. To first order in $\Psi_N$ and $\Phi_N$, the deflection is given as \cite{dodelson2020modern}
\begin{equation}
    \mathbf d(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi),
\end{equation}
where $\Psi$ is the Weyl Potential, $\Psi := (\Psi_N - \Phi_N)/2$, and $\chi_*$ is the conformal distance to the source. $\nabla_{\hat{\mathbf n}}$ is the derivative along the axes orthogonal to the line of sight. The above equation can be written in terms of the lensing potential, $\psi$, as $\mathbf d (\hat{\mathbf n}) = \nabla_{\hat{\mathbf n}} \psi(\hat{\mathbf n})$, with
\begin{equation}
    \psi(\hat {\mathbf n}) := -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
\end{equation}

If the source is instead distributed over radial distance according to some distribution function $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, the $(\chi_* - \chi)/(\chi_*\chi)$ factor is changed as
\begin{equation*}
    \frac{\chi_*-\chi}{\chi_*\chi} \rightarrow W(\chi) := \int_\chi^{\infty}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
\end{equation*}
$W(\chi)$ is then called the window function. In the most general case, the lensing potential is thus given by
\begin{equation}
    \psi(\hat {\mathbf n}) := -2 \int_0^{\infty}\d \chi W(\chi)\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
\end{equation}
The integration limit is sometimes also taken to be the surface of last scattering, as any window function vanishes after that distance. In the case of CMB lensing we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces to $H(\chi_* - \chi)(\chi_*-\chi)/(\chi_*\chi)$, with $H(\chi)$ the Heaviside step function.

@@

\subsection{Convergence and Shear}
The magnification matrix is defined as
\begin{equation}
    A_{ij} := \delta_{ij} + \frac{\partial}{\partial n_j}d_i(\hat{\mathbf n}).
\end{equation}
This matrix can be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
\begin{equation}
    A_{ij}(\hat{\mathbf n}) = \begin{pmatrix}
        1 - \kappa - \gamma_1 & -\gamma_2 + \omega \\
        -\gamma_2 - \omega & 1 - \kappa + \gamma_1
    \end{pmatrix}.
\end{equation}
At 1st order $A$ is a symetric matrix by definition and $\omega$ vanishes, we will ignore it from here on out. Intuitively, $A$ tells you how a small patch in the sky transforms due to lensing. If we change the unlensed direction of a light source by $\delta\hat{\mathbf n}$, then the corresponding change in direction in the lensed image can be calculated as
\begin{equation}
\hat{\mathbf{n}} + \delta \hat{\mathbf{n}} \rightarrow \hat{\mathbf{n}} + \delta \hat{\mathbf{n}} + \mathbf d (\hat{\mathbf{n}} + \delta \hat{\mathbf{n}}) = \hat{\mathbf{n}} + \mathbf d (\hat{\mathbf n}) + \delta \hat{\mathbf n} + A_{ij}(\delta \hat{\mathbf{n}})_j.
\end{equation}
For an image on the sky, $A_{ij}$ thus introduces distortion. Note that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$. We can thus interpret $\kappa$ as telling us about the overal magnification of the source. The $\gamma_i$ represent the area preserving distortion, i.e. stretching and squeezing in a specific direction.

We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
% TODO: should these subscripts be theta and phi?
\begin{equation}
    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_{n_1}^2 - \partial_{n_2}^2)\psi, \quad \gamma_2 = \partial_{n_1}\partial_{n_2} \psi.
\end{equation}

@@

The lensing potential can be decomposed into spherical harmonics as

@@

On the other hand, consider the decomposition of $\Psi$ in Fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,

@@

where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (this is known as the Born approximation), with $\eta_0$ the time at which the light ray hits earth. 

@@

To further evaluate the integral we will make the Limber approximation. The Bessel functions peak sharply at $l=x$\footnote{Some sources use $x\approx l+1/2$ instead, which is slightly more accurate for larger scales (low $l$) and slightly less accurate for smaller scales.}, with the peak being increasingly sharp for higher $l$. Similarly to $\delta(x-x_0)f(x)=\delta(x-x_0)f(x_0)$, we thus take $j_l(k\chi)f(k)\approx j_l(k\chi)f(l/\chi)$. The Bessel functions satisfy an orthogonality condition,

@@

The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is \cite{Hu2000}
\begin{equation*}
    \langle (X_1)_{l_1m_1}(X_2)_{l_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
        l_1 & l_2 & l_3 \\ m_1 & m_2 & m_3
    \end{pmatrix}
    B_{l_1l_2l_3}^{X_1X_2X_3}.
\end{equation*}
Note the independence on $m_i$, this necesarilly follows from statistical isotropy. % TODO: add a short explanation here
If $m_1+m_2+m_3\neq 0$, the associated Wigner-3j symbol vanishes and the bispectrum is set to zero.
Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$. This is relevant when we look at cross correlations between CMB and galaxy lensing.
6758
Using the above definition and the symmetry property
\begin{equation*}
    \begin{pmatrix}
        l_1 & l_2 & l_3\\
        -m_1 & -m_2 & -m_3
      \end{pmatrix}
      =
      (-1)^{l_1+l_2+l_3}
      \begin{pmatrix}
        l_1 & l_2 & l_3\\
        m_1 & m_2 & m_3
      \end{pmatrix},
\end{equation*}
we find
\begin{gather*}
    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
    = (-1)^{l_1+l_2+l_3}
    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
    (2\pi)^3 8 \\
    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \{\eta_0-\chi\}),
\end{gather*}
where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispecrum vanishing if that sum doesn't equal $0$, as mentioned earlier. When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
In particular, it vanishes if $l_1+l_2+l_3$ is odd, meaning we can drop the $(-1)^{l_1+l_2+l_3}$ factor. Additionally cancelling common factors then gives
\begin{align}
    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
    =
    -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi). \label{eq:lbsintermsofpsi}
\end{align}
